---
title: "Data Science with R Assignment"
author: "FAITH LEBONE MOTSHABI"
date: "18/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Basics 
## 1.1 Variables in R

What is the value of " B " after running these command?
```{r}
a <- 5.0
c <- 10
t <- 12
B <- a+c/t
```
The value of B: `r round(B, 2)`

What type of vector do you get in each of the following?
```{r}
vec_1 <- c(44, 5, 0, "k")
vec_2 <- c(1, 2, 3, FALSE)
vec_3 <- c("a", "b", "c", TRUE)
vec_4 <- c(1, 2, 3, "4")
```
The type of vectors:

vec_1: `r typeof(vec_1)`

vec_2: `r typeof(vec_2)`

vec_3: `r typeof(vec_3)`

vec_4: `r typeof(vec_4)`

## 1.2 Conditional subsetting

1.2. Conditional subsetting
Consider the following vector:
```{r}

vec <- c(141.2741, 141.8392, 140.5730,NA, 141.3571, 139.8689, 
         137.9384, NA,139.6620, 138.2521, 138.8635, 141.3885, NA)

vec_1 <- subset(vec, vec != "NULL")
```


1.2.1. Create a vector vec_1 with no missing values using conditional subsetting and explain your code
         
The vec_1 with no missing values: `r vec_1`

Explanation: The function subset() is used to create a subset, vec_1, from the set, vec.In the function the condition is that vec should be printed out but without the NULL values.   

1.2.2. Create a vector of indexes of missing values using conditional subsetting and explain your code

```{r}
vec <- c(141.2741, 141.8392, 140.5730,NA, 141.3571, 139.8689, 
         137.9384, NA,139.6620, 138.2521, 138.8635, 141.3885, NA)

```


1.2.3. Write a code to determine how many missing values are in vec.

```{r}
vec <- c(141.2741, 141.8392, 140.5730,NA, 141.3571, 139.8689, 
         137.9384, NA,139.6620, 138.2521, 138.8635, 141.3885, NA)

```
The number of missing values are: `r sum(is.na(vec))`


1.2.4. Write a R code to create a data frame which contains details of 3 students as follows:
Name
Surname
Age
Gender

```{r}
Students = data.frame(Name=c("Faith","Lebone","Thabang"), Surname=c("Smith", "Motshabi", "Mokwena"), Age=c(20,19,22),Gender=c("Female","Female","Male"))

print("The Students' data frame:")
print(Students)
```
# 2. Functions in R

## 2.1. Create a function which returns a mean of an input vector and explain your code

```{r}
vector <- readline(prompt="Enter number: ")
print(paste(vector))
```

## 2.2. Create a function which take no inputs , it reads a Titanic .csv file and returns its content and explain your code.

```{r}
library(tidyverse)

df <- read.csv("Titanic.csv")

```
Explanation: In the code, the tidyverse library is used to start the project. The csv file is read using read.csv(). The csv file is assigned to df.


# 3. Conditionals and Loops

## 3.1. Write a function that take an input below and returns a transformed vector below into 1 when its element is male and 2 when its female using ifelse.

```{r}
g <- as.factor( c("MALE","FEMALE","FEMALE","MALE","FEMALE","MALE","MALE","FEMALE", "MALE","MALE","FEMALE","MALE","MALE",
"FEMALE","MALE","MALE","MALE"))

g_1 <- ifelse(g == "MALE", "FEMALE", g)

```

The transformed vector: `r g_1`

## 3.2. Generate x from normal distribution and sample with replacement "M" and "F" for male and female respectively by running the code below.

```{r}
x <- rnorm(1000, 120, 5)
y <- sample(c("M","F"),size = 1000, replace = TRUE)

png(file = "rnorm.png")

hist(x, main = "Normal DIstribution")

dev.off()
plot(hist(x, main = "Normal DIstribution"))
```


## 3.2.1. Use the "cbind " function in R to create one dataset and assign it to df

```{r}
x <- rnorm(1000, 120, 5)
y <- sample(c("M","F"),size = 1000, replace = TRUE)
df <- cbind(x,y)
df
```


### 3.2.2. Convert df from 3.2.1 into a data.frame
```{r}
x <- rnorm(1000, 120, 5)
y <- sample(c("M","F"),size = 1000, replace = TRUE)
df <- data.frame(x)
df
```


### 3.2.3. Write a code using tidyverse to compute a number of males ("M") and females ("F") in 3.2.1
```{r}
library(tidyverse)
x <- rnorm(1000, 120, 5)
y <- sample(c("M","F"),size = 1000, replace = TRUE)
df <- cbind(x,y)
as.data.frame(df)
```







